Laplacian envelope

### 2020-10-28 Notes/laplace.tex
copied proof of init bound from older notes

We will be trying to store patterns in a set od $N$ synaptic weights, $\syn$.
Every time we try to store a pattern, these synapses are subjected to a plasticity event where each synapse is either potentiated or depressed, depending on the pattern.
we will assume that these patterns are spatially and temporally independent.

At some time, suppose we wish to determine if a given pattern is one of those that we previously attempted to store.
We wish to answer this question by looking at the synaptic weights directly (ideal observer).
For that given pattern there will be an ideal set of synaptic weights, $\synid$, where those synapses that were supposed to be potentiated are maximised and those that were supposed to be depressed are minimised.
Suppose that the given pattern was actually seen at time 0 and we are observing the synapses at time $t$.
The actual set of synaptic weights we see, $\syn(t)$, will be a vector of random variables that differs from $\synid$ due to the stochasticity of the pattern encoding and all of the other (uncorrelated) pattern that are stored after it.
As $t\to\infty$, the synaptic weights will become independent of the patter stored at $t=0$.
Thus, the vector of random variables $\syn(\infty)$ also describes the synaptic weights under the null hypothesis -- if that given pattern had never been stored.

We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold~\cite{Sommer1998retrieval}.
For large $N$, this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.

@@

The formulae above assumed that we know the time between storage and recognition.
We should also compute the expectations and variances over probability distribution of the recall time, $t$, as well.
If we only know average time, $\tau$, a natural choice for this distribution is
%
\begin{equation}\label{eq:recogtime}
  P(t\vert\tau) = \frac{\e^{-t/\tau}}{\tau}.
\end{equation}
%
Different parts of the brain, that store memories for different timescales, could be characterised by different values of $\tau$.

@@

We will model the synapses as having two possible synaptic weights.
As we are looking at the information contained in the synaptic weights, rather than modelling the readout via electrical activity of the neurons, what values the synaptic weights actually take is irrelevant.
It will be convenient to call these two values $\pm1$.

As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)}}
     {\sqrt{\var(\synid\cdt\syn(\infty))}},
\end{equation}
%
Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
If we also average over the time between storage and recognition, we have
%
\begin{equation}\label{eq:snrb}
  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t)
   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),
\end{equation}
%
This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
%
\begin{equation}\label{eq:laplacedef}
  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t),
  \qquad
  \snrb(\tau) = \frac{A(1/\tau)}{\tau}.
\end{equation}
%

First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\synid\cdt\syn(t))
    &= \sum_{\alpha\beta} \av{\synid^\alpha \syn^\alpha(t) \synid^\beta \syn^\beta(t)}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2 \\
    &= \sum_{\alpha} \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}
    + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
    &= N\av{1}
    + N(N-1)\av{\synid^1 \syn^1(t)}^2
    - N^2\av{\synid^1 \syn^1(t)}^2 \\
    &= N(1-\av{\synid^1 \syn^1(t)}^2),
\end{aligned}
\end{equation}
%
where we used $\syn^\alpha=\pm1$.
We can compute $\var(\synid\cdt\syn(\infty))$ by taking $t\to\infty$.

For the numerator, we can write
%
\begin{equation}\label{eq:overlapex}
  \av{\synid\cdt\syn(t)} = \sum_\alpha\av{\synid^\alpha\syn^\alpha(t)}
   = N \av{\synid^1\syn^1(t)},
\end{equation}
%
If the elements of $\synid$ take values $\pm1$ with probability $f\potdep$,
%
\begin{equation}\label{eq:overlapone}
  \av{\synid^1\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0}
\end{equation}
%
To compute this quantity, we will need to discuss how we model individual synapses.

@@

We model a synapse as having $M$ internal states.
The synaptic weight depends deterministically on the state, given by the $M$-element column vector $\w$.
We denote the probability distribution across these states by the row vector $\pr(t)$.
We denote a column vector of ones by $\onev$, so the normalisation condition is $\pr(t) \, \onev=1$.

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\MMd^\nu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$, 
and \(\nu \) labels the diffrerent types of plasticity event.
Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
We also use the notation
%
\begin{equation*}
  (-1)^\nu = 
    \begin{cases}
       1, & \nu = \text{pot}, \\
      -1, & \nu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices $\MMd^\nu$ are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \MMdm^\nu_{ij} \geq 0 
    \quad \forall \; \nu, i, j, \qquad
  \sum_j \MMdm^\nu_{ij} = 1
    \quad \forall \; \nu, i.
\end{equation}
%
The upper bounds $\MMdm^\nu_{ij} \leq 1$ follow automatically from these.

@@

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$, 
mariginalising over states at earlier times, is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \MMd \pot \MMd \dep \MMd \dep \MMd \pot \MMd \dep \MMd \pot \MMd \pot \MMd \dep \ldots \MMd \pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating or depressing is $f\pot$ or $f\dep$ respectively, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at $t=0$ is fixed, but all of the previous and subsequent events are marginalised over.
If we marginalise over the types (potentiating/depressing) of these other events (conditioned on the number of events) we find
%
\begin{multline}\label{eq:plastseqtypes}
  \pr(t_n) = \pr(t_0) (f\pot \MMd \pot + f\dep \MMd \dep)  \ldots \\
    \ldots (f\pot \MMd \pot + f\dep \MMd \dep) 
     \MMd \potdep (f\pot \MMd \pot + f\dep \MMd \dep) \ldots \\
    \ldots (f\pot \MMd \pot + f\dep \MMd \dep),
\end{multline}
% 
where the factor of $ \MMd \potdep$ is the event at $t=0$.
If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.

If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot \MMd \pot+f\dep \MMd \dep)$:
%
\begin{equation}\label{eq:plastseqeq}
  \pr(t_n) = \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n,
\end{equation}
%
where $n$ is the number of plasticity events that took place.

If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
If we marginalise over $n$ as well
%
\begin{equation}\label{eq:plastseqav}
\begin{aligned}
  P(\text{state}=i,t \mid \text{pot/dep},0)
    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n \\
    &= \brk{ \eq  \MMd \potdep\, \e^{rt(f\pot \MMd \pot+f\dep \MMd \dep-\I) } }_i,
\end{aligned}
\end{equation}
%
where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
%
\begin{equation}\label{eq:eq}
  \eq\frg = 0,
  \qquad\text{where}\quad
  \frg = \sum_\nu f^\nu \MMd^\nu - \I
    = f\pot \MMd \pot + f\dep \MMd \dep - \I.
\end{equation}
%

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq \w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq \w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\MMd^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\MMdm^\mu_{ij}}{\MMdm^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMd \dep_{ij}},
  \qquad
  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frgm_{ij}}{\frgm_{mn}} = \pdiff{\encm_{ij}}{\encm_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frgm_{ij}}{\encm_{mn}} = \pdiff{\encm_{ij}}{\frgm_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq f^\mu, \\
  \sum_j \frgm_{ij} &=0, &
  \sum_j \encm_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

In analogy to the generalised fundamental matrix of a Markov chain \cite{Kemeny1981fund}, define
%
\begin{equation}\label{eq:lfund}
  \fund(s) = (s \I + \onev \arow -\MM)\inv,
\end{equation}
%
where $\arow$ is an arbitrary row vector satisfying $\arow \onev = \frac{1}{\tau_0} \neq 0$.
This reduces to the fundamental matrix at $s=0$.
We can see that $\onev$ is an eigenvector:
%
\begin{equation}\label{eq:lfundrowsum}
  \fund(s)\onev = \frac{\tau_0}{1+s\tau_0}\onev.
\end{equation}
%
Suppose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
Then $\mathbf{a}\fund(s)$ is independent of $\arow$:
%
\begin{equation}\label{eq:alfund}
  \pdiff{(\mathbf{a}\fund(s))_i}{\arowm_k} = -(\mathbf{a}\fund(s)\onev) \fundm_{ki}(s)
      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fundm_{ki}(s) = 0.
\end{equation}
%

Then we also define
%
\begin{equation}\label{eq:lfptb}
  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\Eq\inv,
  \qquad
  \fptbm_{ij}(s) = \frac{\fundm_{jj}(s)-\fundm_{ij}(s)}{\eqm_j}.
\end{equation}
%
This reduces to the mean first-passage time matrix at $s=0$.

@@

This means that the \( \fptb(s) \) have all the same properties as in the case $s=0$.
In particular, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptbm_{ij}(s) \eqm_j
\end{equation}
%
is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}$.
We can then define
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptbm_{ij}(s) \eqm_j \prn{\frac{1 \pm \wm_j}{2}} 
      =  \sum_{j\in \pm} \fptbm_{ij}(s) \eqm_j.
\end{equation}
%
We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
It will be more convenient to use the quantities
%
\begin{equation}\label{eq:lwpmixtime}
  \etwm_i(s) = \sum_j \fptbm_{ij}(s)\eqm_j \wm_j
    = \eta^+_i(s) - \eta^-_i(s) 
    = 2\eta^+_i(s) - \eta(s) 
    = \eta(s) - 2\eta^-_i(s) .
\end{equation}
%
Arranging the states in order of decreasing $\etwm_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
The $\etwm_i(s)$ can still be used when $\w$ takes more than two values.

>>>>

We will be trying to store patterns in a set od \(N\) synaptic weights, \(\syn\).
Every time we try to store a pattern, these synapses are subjected to a plasticity event where each synapse is either potentiated or depressed, depending on the pattern.
we will assume that these patterns are spatially and temporally independent.

At some time, suppose we wish to determine if a given pattern is one of those that we previously attempted to store.
We wish to answer this question by looking at the synaptic weights directly (ideal observer).
For that given pattern there will be an ideal set of synaptic weights, \(\synid\), where those synapses that were supposed to be potentiated are maximised and those that were supposed to be depressed are minimised.
Suppose that the given pattern was actually seen at time 0 and we are observing the synapses at time \(t\).
The actual set of synaptic weights we see, \(\syn(t)\), will be a vector of random variables that differs from \(\synid\) due to the stochasticity of the pattern encoding and all of the other (uncorrelated) pattern that are stored after it.
As \(t\to\infty\), the synaptic weights will become independent of the patter stored at \(t=0\).
Thus, the vector of random variables \(\syn(\infty)\) also describes the synaptic weights under the null hypothesis -- if that given pattern had never been stored.

We can test if the pattern had been previously stored by computing \(\synid\cdt\syn\) and comparing it to some threshold~\cite{Sommer1998retrieval}.
For large \(N\), this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, \(\snr\) is larger than \(\Phi_c\inv(\operatorname{FPR})\), it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, \(\synid\).

@@

The formulae above assumed that we know the time between storage and recognition.
We should also compute the expectations and variances over probability distribution of the recall time, \(t\), as well.
If we only know average time, \(\tau\), a natural choice for this distribution is
%
\begin{equation}\label{eq:recogtime}
  P(t\vert\tau) = \frac{\e^{-t/\tau}}{\tau}.
\end{equation}
%
Different parts of the brain, that store memories for different timescales, could be characterised by different values of \(\tau\).

@@

We will model the synapses as having two possible synaptic weights.
As we are looking at the information contained in the synaptic weights, rather than modelling the readout via electrical activity of the neurons, what values the synaptic weights actually take is irrelevant.
It will be convenient to call these two values \(\pm1\).

As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)}}
     {\sqrt{\var(\synid\cdt\syn(\infty))}},
\end{equation}
%
Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, \(\synid\).
If we also average over the time between storage and recognition, we have
%
\begin{equation}\label{eq:snrb}
  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t)
   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),
\end{equation}
%
This is similar to the average SNR up to time \(\tau\), but with the hard cut-off replaced with a smooth exponential.
This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
%
\begin{equation}\label{eq:laplacedef}
  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t),
  \qquad
  \snrb(\tau) = \frac{A(1/\tau)}{\tau}.
\end{equation}
%

First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\synid\cdt\syn(t))
    &= \sum_{\alpha\beta} \av{\synid^\alpha \syn^\alpha(t) \synid^\beta \syn^\beta(t)}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2 \\
    &= \sum_{\alpha} \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}
    + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
    &= N\av{1}
    + N(N-1)\av{\synid^1 \syn^1(t)}^2
    - N^2\av{\synid^1 \syn^1(t)}^2 \\
    &= N(1-\av{\synid^1 \syn^1(t)}^2),
\end{aligned}
\end{equation}
%
where we used \(\syn^\alpha=\pm1\).
We can compute \(\var(\synid\cdt\syn(\infty))\) by taking \(t\to\infty\).

For the numerator, we can write
%
\begin{equation}\label{eq:overlapex}
  \av{\synid\cdt\syn(t)} = \sum_\alpha\av{\synid^\alpha\syn^\alpha(t)}
   = N \av{\synid^1\syn^1(t)},
\end{equation}
%
If the elements of \(\synid\) take values \(\pm1\) with probability \(f\potdep\),
%
\begin{equation}\label{eq:overlapone}
  \av{\synid^1\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0}
\end{equation}
%
To compute this quantity, we will need to discuss how we model individual synapses.

@@

We model a synapse as having \(M\) internal states.
The synaptic weight depends deterministically on the state, given by the \(M\)-element column vector \(\w\).
We denote the probability distribution across these states by the row vector \(\pr(t)\).
We denote a column vector of ones by \(\onev\), so the normalisation condition is \(\pr(t) \, \onev=1\).

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices \(\MMd^\nu\) whose \(i,j\) elements are the transition probabilities from state \(i\) to state \(j\), 
and \(\nu \) labels the diffrerent types of plasticity event.
Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
We also use the notation
%
\begin{equation*}
  (-1)^\nu = 
    \begin{cases}
       1, & \nu = \text{pot}, \\
      -1, & \nu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices \(\MMd^\nu\) are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \MMdm^\nu_{ij} \geq 0 
    \quad \forall \; \nu, i, j, \qquad
  \sum_j \MMdm^\nu_{ij} = 1
    \quad \forall \; \nu, i.
\end{equation}
%
The upper bounds \(\MMdm^\nu_{ij} \leq 1\) follow automatically from these.

@@

Given a sequence of plasticity events at times \(t_1,t_2,t_3,\ldots\), the probability distribution across the internal states at time \(t_n\), 
mariginalising over states at earlier times, is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \MMd \pot \MMd \dep \MMd \dep \MMd \pot \MMd \dep \MMd \pot \MMd \pot \MMd \dep \ldots \MMd \pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating or depressing is \(f\pot\) or \(f\dep\) respectively, the probability of the sequence above (conditioned on the number of events) is given by \(f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot\).

In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at \(t=0\) is fixed, but all of the previous and subsequent events are marginalised over.
If we marginalise over the types (potentiating/depressing) of these other events (conditioned on the number of events) we find
%
\begin{multline}\label{eq:plastseqtypes}
  \pr(t_n) = \pr(t_0) (f\pot \MMd \pot + f\dep \MMd \dep)  \ldots \\
    \ldots (f\pot \MMd \pot + f\dep \MMd \dep) 
     \MMd \potdep (f\pot \MMd \pot + f\dep \MMd \dep) \ldots \\
    \ldots (f\pot \MMd \pot + f\dep \MMd \dep),
\end{multline}
% 
where the factor of \( \MMd \potdep\) is the event at \(t=0\).
If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.

If \(t_0 \ll 0\), the effect of the large number of earlier events will be to put the synapse in the steady state distribution, \(\eq\), of the stochastic process \((f\pot \MMd \pot+f\dep \MMd \dep)\):
%
\begin{equation}\label{eq:plastseqeq}
  \pr(t_n) = \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n,
\end{equation}
%
where \(n\) is the number of plasticity events that took place.

If the plasticity events occur at Poisson rate \(r\), the probability of \(n\) events occurring in time \(t\) is given by a Poisson distribution with mean \(rt\).
If we marginalise over \(n\) as well
%
\begin{equation}\label{eq:plastseqav}
\begin{aligned}
  P(\text{state}=i,t \mid \text{pot/dep},0)
    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n \\
    &= \brk{ \eq  \MMd \potdep\, \e^{rt(f\pot \MMd \pot+f\dep \MMd \dep-\I) } }_i,
\end{aligned}
\end{equation}
%
where \(\eq\) is the steady state distribution of the continuous time transition matrix that appears in the exponential:
%
\begin{equation}\label{eq:eq}
  \eq\frg = 0,
  \qquad\text{where}\quad
  \frg = \sum_\nu f^\nu \MMd^\nu - \I
    = f\pot \MMd \pot + f\dep \MMd \dep - \I.
\end{equation}
%

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq \w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with \(f\pot=f\dep\) or \(\eq_+=\eq_-\).
We can perform the maximisation in two steps.
First maximise the numerator at fixed \(\eq \w\).
Then maximise the ratio \wrt \(\eq\w\).
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of \(\MMd^\nu\) as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\MMdm^\mu_{ij}}{\MMdm^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMd \dep_{ij}},
  \qquad
  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frgm_{ij}}{\frgm_{mn}} = \pdiff{\encm_{ij}}{\encm_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frgm_{ij}}{\encm_{mn}} = \pdiff{\encm_{ij}}{\frgm_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of \(\frg\) and \(\enc\) as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq f^\mu, \\
  \sum_j \frgm_{ij} &=0, &
  \sum_j \encm_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

In analogy to the generalised fundamental matrix of a Markov chain \cite{Kemeny1981fund}, define
%
\begin{equation}\label{eq:lfund}
  \fund(s) = (s \I + \onev \arow -\MM)\inv,
\end{equation}
%
where \(\arow\) is an arbitrary row vector satisfying \(\arow \onev = \frac{1}{\tau_0} \neq 0\).
This reduces to the fundamental matrix at \(s=0\).
We can see that \(\onev\) is an eigenvector:
%
\begin{equation}\label{eq:lfundrowsum}
  \fund(s)\onev = \frac{\tau_0}{1+s\tau_0}\onev.
\end{equation}
%
Suppose we have some row vector \(\mathbf{a}\) such that \(\mathbf{a}\onev=0\).
Then \(\mathbf{a}\fund(s)\) is independent of \(\arow\):
%
\begin{equation}\label{eq:alfund}
  \pdiff{(\mathbf{a}\fund(s))_i}{\arowm_k} = -(\mathbf{a}\fund(s)\onev) \fundm_{ki}(s)
      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fundm_{ki}(s) = 0.
\end{equation}
%

Then we also define
%
\begin{equation}\label{eq:lfptb}
  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\Eq\inv,
  \qquad
  \fptbm_{ij}(s) = \frac{\fundm_{jj}(s)-\fundm_{ij}(s)}{\eqm_j}.
\end{equation}
%
This reduces to the mean first-passage time matrix at \(s=0\).

@@

This means that the \( \fptb(s) \) have all the same properties as in the case \(s=0\).
In particular, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptbm_{ij}(s) \eqm_j
\end{equation}
%
is independent of the starting state \(i\), but now \(\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}\).
We can then define
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptbm_{ij}(s) \eqm_j \prn{\frac{1 \pm \wm_j}{2}} 
      =  \sum_{j\in \pm} \fptbm_{ij}(s) \eqm_j.
\end{equation}
%
We can arrange the states in order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
It will be more convenient to use the quantities
%
\begin{equation}\label{eq:lwpmixtime}
  \etwm_i(s) = \sum_j \fptbm_{ij}(s)\eqm_j \wm_j
    = \eta^+_i(s) - \eta^-_i(s) 
    = 2\eta^+_i(s) - \eta(s) 
    = \eta(s) - 2\eta^-_i(s) .
\end{equation}
%
Arranging the states in order of decreasing \(\etwm_i(s)\) is the same as the order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
The \(\etwm_i(s)\) can still be used when \(\w\) takes more than two values.
